Evolution of CSMA Protocols for the IEEE 802.11 Standard

### 2013-01-24 paper.tex
enhanced csma/eca
removed r
changed abstract

\begin{abstract}
%\boldmath
In this paper we present the requirements of candidate protocols to replace the pervasive CSMA/CA medium access control.
We discuss the possibility of further preventing collisions and provide an overview of the related work.
We specify protocols that are candidates of replacing CSMA/CA in pseudocode and use simulation to assess performance metrics such as throughput, fairness and collision probability.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Enhanced CSMA}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        $s \leftarrow 0$ \;
        $b \leftarrow \mathcal{U}[0,2^{s}\rm{CW}_{min}-1]$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      $s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0,2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/CA}
\label{alg:CSMA_CA}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        $s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      $s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA}
\label{alg:CSMA_ECA}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        %$s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      %$s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA with hysteresis}
\label{alg:CSMA_ECA_hist}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission of $2^s$ packets \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        %$s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      %$s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA with hysteresis and fair-share}
\label{alg:CSMA_ECA_hist_fair}
\end{algorithm}

>>>>

\begin{abstract}
%\boldmath
In this paper we present the requirements of candidate protocols to replace the pervasive CSMA/CA medium access control.
We discuss the possibility of further preventing collisions and provide an overview of the related work.
We specify a protocol that is a candidate for replacing CSMA/CA in pseudocode and use simulation to assess performance metrics such as throughput, fairness and collision probability.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Enhanced CSMA/CA}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        $s \leftarrow 0$ \;
        $b \leftarrow \mathcal{U}[0,2^{s}\rm{CW}_{min}-1]$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min (s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      $s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0,2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/CA}
\label{alg:CSMA_CA}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        $s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min (s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      $s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA}
\label{alg:CSMA_ECA}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        %$s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min (s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      %$s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA with hysteresis}
\label{alg:CSMA_ECA_hist}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission of $2^s$ packets \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        %$s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min (s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      %$s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA with hysteresis and fair-share}
\label{alg:CSMA_ECA_hist_fair}
\end{algorithm}
